---
title: "Linear Models"
format: html
editor: visual
---

## Setup
```{r}
library(tidyverse)
library(corrplot)
library(car)

df <- read_csv("merged_data.csv")
```


## Model Ideas

#### What we need to analyze
- Better football season -> Enrollment
  - Within a team, avg test scores two years later ~ percentage of games won
  - Within a team, avg enrollment rates two years later ~ percentage of games won
- Better football season -> Male/female ratio
  - Within a team, male ratio two years later ~ percentage of games won
- Better football season -> Racial Diversity
  - Within a team, white ratio two years later ~ percentage of games won
  - Within a team, hispanic ratio two years later ~ percentage of games won
  - Within a team, asian ratio two years later ~ percentage of games won




#### Get data on teams

```{r}
# Using Alabama data
data <- df[c(1:21),]

# Using Boston College data
data <- df[c(200:220),]

# Using a Random Team's data
set.seed(24601)
# Find Team ID
team_ids <- unique(df$UNITID)
rand_int <- sample(1:112, 1)
rand_id <- team_ids[rand_int]
# Get Data
data <- subset(df, UNITID == rand_id)
print(data$team[1])
```

#### Avg SAT scores two years later ~ % games won
```{r}
# Find x
x <- data$won / (data$won + data$lost)
x <- x[1:(length(x)-2)]

# Find y
y <- data$SAT_AVG
y <- y[3:length(y)]

mod <- lm(y ~ x)

# Assumptions
plot(mod)

summary(mod)
plot(x, y)
```

#### Avg addmission rates two years later ~ % games won
```{r}
# Find x
x <- data$won / (data$won + data$lost)
x <- x[1:(length(x)-2)]

# Find y
y <- data$ADM_RATE
y <- y[3:length(y)]

mod <- lm(y ~ x)

# Assumptions
plot(mod)

summary(mod)
plot(x, y)
```

#### Within a team, male ratio two years later ~ percentage of games won
```{r}

```

#### Within a team, white ratio two years later ~ percentage of games won
```{r}
# Find x
x <- data$won / (data$won + data$lost)
x <- x[1:(length(x)-2)]

# Find y
y <- data$UGDS_WHITE
y <- y[3:length(y)]

mod <- lm(y ~ x)

# Assumptions
plot(mod)

summary(mod)
plot(x, y)
```

#### Within a team, hispanic ratio two years later ~ percentage of games won
```{r}
# Find x
x <- data$won / (data$won + data$lost)
x <- x[1:(length(x)-2)]

# Find y
y <- data$UGDS_HISP
y <- y[3:length(y)]

mod <- lm(y ~ x)

# Assumptions
plot(mod)

summary(mod)
plot(x, y)
```

#### Within a team, asian ratio two years later ~ percentage of games won
```{r}
# Find x
x <- data$won / (data$won + data$lost)
x <- x[1:(length(x)-2)]

# Find y
y <- data$UGDS_ASIAN
y <- y[3:length(y)]

mod <- lm(y ~ x)

# Assumptions
plot(mod)

summary(mod)
plot(x, y)
```








